\newpage
\section{Transformation Rule for TensorFlow ML Training Code}
\subsection{Restrictions}
\begin{itemize}
  \item[1.] All import statements must be placed at the top of a module.  
  \item[2.] The {\tt tensorflow} module must be assigned to a variable only
    using an import statement.
  \item[3.] Members of the {\tt tensorflow} module must not be aliased.
  \item[4.] {\tt print} function call expressions must not introduce
    side-effects.
  \item[5.] A dataset and an optimizer object must only be created once via an
    assignment statement with a function call expression and must not be
    aliased.
  \item[6.] Variables storing a dataset or a optimizer object must not be
    reassigned to store another value that is not a dataset or an optimizer
    object.
  \item[7.] A dataset and an optimizer object must not be created
    conditionally.
  \item[8.] {\tt optimizer.apply\_gradients} function call expressions must be
    expression statements or a direct RHS of an assignment statement.
  \item[9.] Global variables storing an {\tt optimizer} object and referred in
    functions must be defined before the functions' definitions and must not be
    changed after their initializations.
  \item[10.] A checkpoint object must only be created once via an assignment
    statement with a function call expression and must not be aliased. 
\end{itemize}
\subsection{Rules}
\subsubsection{Types and Auxiliary Functions}
% Domains
\noindent
\begin{tabular}{lcl>{\sc(}l<{\sc)}}
  $\tau$ & ::= & \dmodule & Modules\\
  &&\dstmt & Statements\\
  &&\dexpr & Expressions\\
  &&\dcomp & Comprehensions\\
  &&\dhandler& Exception handlers\\
  &&\dalias & Aliases\\
  &&\dwithitem & Withitems\\
  &&\dpattern & Patterns\\
  &&\did& Identifiers\\
  &&\dstr& Strings\\
  &&$\tau$ list& List of elements typed $\tau$\\
\end{tabular}\\\vpar

\noindent
\begin{tabular}{l}
  \smodenv ~ $\in$ \dmodenv ~ = \dstr $\finarrow$ \did \qquad {\sc Environment storing mappings from strings to identifiers}\\
\end{tabular}\\\vpar

\noindent
\begin{tabular}{lcl>{\sc}l}
  \fst & : & $\tau \times \tau \rightarrow \tau$ & Get the first element of the given pair\\
  \ktconl & : & $\tau$ list $\rightarrow \tau$ list $\rightarrow \tau$ list & Concatenate two lists\\
  \ktappl & : & $\tau$ $\rightarrow \tau$ list $\rightarrow \tau$ list & Append an element to a list (right-associative)\\
  \ktsubty & : & \dexpr $\times$ \dexpr $\rightarrow$ $\mathbb{B}$ & Whether classes specified by the expressions are subclasses \\
\end{tabular}\\\vpar
