\documentclass{beamer}
\usetheme{Oxygen}
\usepackage{verbatim}

\title{Automated Code Transformation for Distributed Training of TensorFlow ML Models}
\author{Yusung Sim\inst{1}, Wonho Shin\inst{1}, Sungho Lee\inst{2}}
\institute{
  \inst{1}%
  School of Computing, KAIST
  \and
  \inst{2}%
  Department of Computer Science and Engineering, Chungnam National University
}
\date{2022}

\begin{document}
% 0. Title
\frame{\titlepage}


% 1. Introduction
\begin{frame}
  \frametitle{Distributed ML Training}
  Distributed ML Training scales well \& reduces training time.
\end{frame}

\begin{frame}
  \frametitle{Challenge: Distributing single-GPU based model}
  Rewriting single-GPU based model into distributed model
  \begin{itemize}
    \item Manual rewriting is tedious and time-consuming
    \item Developer must understand distributed training API
    \item Actual code changes are mostly syntactic and simple
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Automatic code transformation}
  Propose automatic code transformation for distributed ML training
  \begin{itemize}
    \item Define formal transformation rule
    \item Implement the rule as a software
  \end{itemize}
\end{frame}

% 2. Background
\begin{frame}
  \frametitle{Horovod: Distributed Training Framework}
  Horovod is a distributed ML training framework developed by Uber, 2017.
\end{frame}

% 3. Method
\begin{frame}
  \frametitle{Python Abstract Syntax}
  We manually investigated the Python Language Reference,
  defined the Python abstract syntax.
\end{frame}

\begin{frame}
  \frametitle{Training API Patterns for TensorFlow ML Model}
  TensorFlow provides multiple training APIs.
  Different training API usage requires different transformation rule 

\end{frame}

\begin{frame}
  \frametitle{Training API Patterns for TensorFlow ML Model}
  We defined 4 API Patterns
\end{frame}

\begin{frame}
  \frametitle{Formal Transformation Rule for Distributed ML Model}
  Code transformation is formally defined as transform function
  AST $\rightarrow$ AST
  
  Env parameter stores specific identifiers and passes to the next call
  
\end{frame}

\begin{frame}
  \frametitle{Formal Transformation Rule for Distributed ML Model}
  Explain the application of tramsform function by example  
\end{frame}


% 4. Implementation
\begin{frame}
  \frametitle{Distributed ML Code Transformer Software}
  We implemented the transformation rule as a software.
  Scala
\end{frame}

\begin{frame}
  \frametitle{Impl. Class Hierarchy Analysis}
\end{frame}

\begin{frame}
  \frametitle{Impl. Training API Pattern Analysis}
\end{frame}

\begin{frame}
  \frametitle{Impl. Transformer Function}
\end{frame}

% 5. Evaluation
\begin{frame}
  \frametitle{}
\end{frame}

% for record
\begin{frame}[fragile]
  \frametitle{Code example}
  \begin{verbatim}
  def add(x, y):
    return x + y
  \end{verbatim}
\end{frame}

\end{document}
